The Driving Forces of API Evolution

### 2015-06-03 thesis/paper.tex
Further minor changes

The API analysed in this study is written in the programming language C and has been deployed for several years. It was recently updated from v1.4 to v2.0, which is the current active version. The changes between these two versions are what has been analysed in this study. Between the release of v1.0 and v1.4, no functionality was added, as it only added additional build options for other hardware platforms. Prior to v1.4 being released, updates were only made to include additional build options and did not include any additional functionality. The API is used by company partners, which are other companies developing applications for the cameras. These applications are, in turn, used by the end-users of the cameras. The API is compatible with a wide range of camera types that offer different functionality and are used for different purposes.

\subsubsection{APIs as Systems}
The API is rather complex, and in addition to the API interface and implementation code, it includes a middle-layer which manages error handling and the access to various 3\textsuperscript{rd} party libraries. In this study, we use the API interface and the middle-layers as units of analysis. Examples of where the responsibilites exceed those of a primitive API interface is that it includes functionality such as error handling and a daemon which facilitates communication between hardware entities. In addition to that, the SDK which includes the API documentation, user manuals and example applications that utilise the API code. Based on these characteristics and the IEEE definition of a system, which reads as follows; ``A combination of interacting elements organised to achieve one or more stated purposes...Organised to accomplish a specific function or set of functions within a specific environment" \cite{5733835 }, we consider the API to be a system because of the combination of the interacting parts of the API, coupled with the specific nature of the problem the API is addressing.

Building on this, we regard the API examined in this study to be an E-type system based on that it is a part of the world it models. This is evident from that it is a vital part of the cameras' and the only option for developing software for the cameras. The API is also by company A considered to be a vital part of the cameras' software ecosystem. This factor also corresponds with the characteristic of E-type systems which relates to the system being susceptible to change from pressure that occurs after the system has been deployed. Similarly, we do not consider the API to be a P-type system, based on that the system is virtually impossible to model because of its high dependenace with the environment it operates in.

@@

\smallskip \noindent
\textbf{Construct validity  } Since our study is largely based on existing theories \cite{chapin2001types, lehman90sview}, the validity of it directly correlates to the validity of the theories and the applicability of them. With regards to Lehman's laws of software evolution, the laws are said to only apply to E-type systems. According to our analysis, we consider the API to be of the E-type, but our interpretations of Lehman's definitions are incorrect, it could be considered a threat to the construct of our study.

\smallskip \noindent
\textbf{Internal validity  } Limitations related to internal validity have been acknowledged by analysing the code and documentation jointly. This has contributed to triangulating the results and discovering possible inconsistencies. Further triangulation of the results was performed by interviewing the API architect. Business-related factors might have had an impact on the evolution of the API, and these were not closely investigated. This was not in the scope of the study and would best be investigated in a study complementary to ours. The study revealed that the API users were significant to the evolution of the API. Since API users were not interviewed, it might be an affecting factor which was left unexplored.

\smallskip \noindent
\textbf{External validity  } Since only two versions of the API were analysed, we cannot claim any generalisability for our study. We do, however, believe that our study can offer comparability for companies that chose to do a similar analysis on their API.

@@

In total, thirteen unique changes were identified when analysing the source code, the majority of which were additions of functionality of the API. Such additions include added functionality to allow mechanical and digital control of the camera, to utilise additional storage devices, to allow serial connectivity and to allow audio analysis on the cameras. Changes that were pure additions the API were not attributed with any secondary types of change.

Many of the changes were deprecations that were later re-implemented as new modules. Changes of these kinds were grouped into a single type instead of two. Such examples include the deprecation and re-implementation of an event system, dynamic web page generation and configuration utilities. Changes of this kind often had an accompanying secondary type of change.

@@

%Types intro
The results show that certain types of change are more prevalent than others, these results are shown in Table \ref{table:type_counts}. Analysis of the results showed that enhancive changes were the primary types of change that drive API evolution and that they act as a catalyst for many of the other changes in the API. Updative changes were found to be consequences of enhancive changes, rather than being drivers of evolution. Supporting the users of the API through the combination of preventive and groomative types of change, together formed another strong driver of API evolution. Reductive and adaptive changes, while important, did not strongly drive the evolution of the API. No evidence was found for corrective or performance types of change. Similarly, there was no evidence found for evaluative, consultive or training activities.

>>>>

The API analysed in this study is written in the programming language C and has been deployed for several years. It was recently updated from v1.4 to v2.0, which is the current active version. The changes between these two versions are what has been analysed in this study. Between the release of v1.0 and v1.4, no functionality was added, as it only added additional build options for other hardware platforms. The API is used by company partners, which are other companies developing applications for the cameras. These applications are, in turn, used by the end-users of the cameras. The API is compatible with a wide range of camera types that offer different functionality and are used for different purposes.

\subsubsection{APIs as Systems}
Company A's API is rather complex, and in addition to the API interface and implementation code, it includes a middle-layer which manages error handling and the access to various 3\textsuperscript{rd} party libraries. In this study, we use the API interface and the middle-layers as units of analysis. Examples of where the responsibilities exceed those of a primitive API interface is that it includes functionality such as error handling and a daemon which facilitates communication between hardware entities. In addition to that is an SDK which includes the API documentation, user manuals and example applications that utilise the API code. Based on these characteristics and the IEEE definition of a system, which reads as follows; ``A combination of interacting elements organised to achieve one or more stated purposes...Organised to accomplish a specific function or set of functions within a specific environment" \cite{5733835 }, we consider the API to be a system because of the combination of the interacting parts of the API, coupled with the specific nature of the problem the API is addressing.

Building on this, we regard the API examined in this study to be an E-type system based on that it is a part of the world it models. This is evident from that it is a vital part of the cameras' and the only option for developing software for the cameras. The API is also considered by company A to be a vital part of the cameras' software ecosystem. This factor corresponds with the characteristic of E-type systems which relates to the system being susceptible to change from pressure that occurs after the system has been deployed. Similarly, we do not consider the API to be a P-type system, based on the fact that the system is virtually impossible to model due to its high dependence on the environment in which it operates in.

@@

\smallskip \noindent
\textbf{Construct validity  } Since our study is largely based on existing theories \cite{chapin2001types, lehman90sview}, the validity of it directly correlates to the validity of the theories and the applicability of them. With regards to Lehman's laws of software evolution, the laws are said to only apply to E-type systems. According to our analysis, we consider the API to be of the E-type, but if our interpretations of Lehman's definitions are incorrect, it could be considered a threat to the construct validity of our study.

\smallskip \noindent
\textbf{Internal validity  } Limitations related to internal validity have been acknowledged by analysing the code and documentation jointly. This has contributed to triangulating the results and discovering possible inconsistencies. Further triangulation of the results was performed by interviewing an architect of the API. Business-related factors might have had an impact on the evolution of the API, and these were not closely investigated. This was not in the scope of the study and would best be investigated in a study complementary to ours. The study revealed that the API users were significant to the evolution of the API. Since API users were not interviewed, it might be an affecting factor which was left unexplored.

\smallskip \noindent
\textbf{External validity  } Since only two versions of the API were analysed, we cannot claim any generalisability for our study. We do, however, believe that our study can offer comparability for companies that chose to do a similar analysis on their own APIs.

@@

In total, thirteen unique changes were identified when analysing the source code, the majority of which were additions of functionality of the API. Such additions include added functionality to allow mechanical and digital control of the camera, to utilise additional storage devices, to allow serial connectivity and to allow audio analysis on the cameras. Changes that were pure additions to the API were not attributed with any secondary types of change.

Many of the changes were deprecations that were later re-implemented as new modules. Changes of this kind were grouped into a single type instead of two. Such examples include the deprecation and re-implementation of an event system, dynamic web page generation and configuration utilities. Changes of this kind often had an accompanying secondary type of change.

@@

%Types intro
The results show that certain types of change are more prevalent than others, these results are shown in Table \ref{table:type_counts}. Analysis of the results showed that enhancive changes were the primary types of change that drove the evolution of this API and that they acted as a catalyst for many of the other changes in the API. Updative changes were found to be consequences of enhancive changes, rather than being drivers of evolution themselves. Supporting the users of the API through the combination of preventive and groomative types of change, together formed another strong driver of the API's evolution. Reductive and adaptive changes, while important, did not strongly drive the evolution of the API. No evidence was found for corrective or performance types of change. Similarly, there was no evidence found for evaluative, consultive or training activities.
